### 2024-11-30 00_JRR-submission-folder_elsevier-latex-template_COMPLETED/cas-dc-template.tex
update

%% 
%% Copyright 2019-2024 Elsevier Ltd
%% 
%% This file is part of the 'CAS Bundle'.
%% --------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.3c of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.3c or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'CAS Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for cas-dc documentclass for 
%% double column output.

\documentclass[a4paper,fleqn]{cas-dc}

@@

%\documentclass[a4paper,fleqn,longmktitle]{cas-dc}

%\usepackage[numbers]{natbib}
%\usepackage[authoryear]{natbib}
\usepackage[authoryear,longnamesfirst]{natbib}

@@

% Uncomment and use as if needed
%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}[theorem]{Lemma}
%\newdefinition{rmk}{Remark}
%\newproof{pf}{Proof}
%\newproof{pot}{Proof of Theorem \ref{thm}}

@@

% Abstract
\begin{abstract}

%? reviewer 2 - comment 2 - simplify language
The quintessential principle of the circular economy is to minimise material consumption and waste generation. Thus, identifying and quantifying waste and material flows is critical. Life Cycle Assessment (LCA) is a powerful method, given its capacity to pinpoint hotspots of environmental impact throughout an activity's life cycle, where the implementation of circular principles could be most effective.

We introduce T-reX, a Python tool extending the brightway ecosystem to allow facile quantification of user-defined supply chain demands in current and prospective scenarios. T-reX streamlines database manipulation for LCA practitioners and integrates methods to aggregate and analyse waste and material flows, facilitating rapid hotspot identification.

A case study of five lithium-ion batteries demonstrates T-reX's utility, quantifying categorised waste and material inventory footprints and their potential environmental burdens. T-reX can aid sustainable decision-making and contribute to the development of a `circular economy' by facilitating analysis of material consumption and waste generation in LCA.

@@

% Keywords
\verb+\begin{keyword}+ \verb+circular economy+ \verb+waste+ \verb+material+ \verb+life cycle assessment+ \verb+critical raw material+ \verb+supply chain+ \end{keyword}

% Graphical abstract
\begin{graphicalabstract}
  \includegraphics{figs/cas-grabs-TREX.pdf}
\end{graphicalabstract}

@@

% Keywords
\begin{keyword}
  circular economy \sep{}waste \sep{}material \sep{}life cycle assessment \sep{}critical raw material \sep{}supply chain
\end{keyword}

@@

%? reviewer 2 - comment 2 - simplify language
Development of a circular economy has become a central focus in the pursuit of sustainability objectives toward curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}. Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{reike2018rex, eu2022ecodesign, eu2022repair, eu2015reman}. In addition to circular economy goals, contemporary geopolitical tensions in an ever more globalised economy have highlighted the vulnerability of many advanced economies to intentional supply disruptions---wrought as an act of competition or even outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.

@@

%? reviewer 2 - comment 1 - COMPARISON OF T-REX TO OTHER TOOLS
%! fix
Life Cycle Assessment (LCA) is a useful method for the holistic estimation of the environmental impacts of products and processes. LCA can comprehensively evaluate impacts across the entire life cycle---from `cradle to grave'--- often identifying critical hotspots and guiding prioritisation of action~\citep{guinee2011lca}. The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe} and CML~\citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows (those between the technosphere and biosphere). These scores can then be aggregated into one metric for each impact category that can be compared across products and processes. Extending on standard LCA is \textit{ex-ante} LCA, which employs future scenarios to construct `prospective background databases' to predict the impacts of supply chains that have not yet been (and may never be) realised~\citep{cucurachi2018exante,blanco2020exante}.

@@

Following the logic of the standard LCA models \cite{ecoinvent2016version3, guinee2004economicallocation}, the logic for the identification of material exchanges with T-reX differs from that used to identify waste exchanges in that the search queries are based on the names of the relevant `market activities' for the material of interest rather than keywords in the exchange's name. A useful feature of T-reX is that, in cases where there are several markets for one material or material group, the program can aggregate these flows. For example, exchanges with markets for the rare-earth-elements (REEs) `market for cerium', `market for dysprosium', `market for erbium', etc.\ can be aggregated into a single indicator category for REEs. Similarly, the total demand for all critical raw materials (CRMs) can be calculated in the same manner.

@@

\subsubsection{LCA calculations}
For each combination of activity, method, and database, a total summation was calculated along with details of the top contributing processes. Additionally, for the T-reX methods, a contribution analysis was performed with the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, a component of the \texttt{brightway} ecosystem. This function performs one subjective method of graph traversal on the impact matrix of the LCA object to a specified cutoff. The groups of the resulting leaves are collected by their Cooperative Patent Classification (CPC) codes [giving a representation of industry sector involvement]. T-reX provides, essentially, a tabulation of internal exchanges and an insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand footprints.

@@

% Back matter
\section*{Data availability}
All data used is publically available online under the noted sources. T-reX is fully open source and dedicated to the public domain under the CC-0 licence. It is installable via the Python Package Index (PyPI) under the name `T\_reX\_LCA' \url{https://pypi.org/project/T_reX_LCA}.
The full source code for T-reX is archived by Zenodo at \url{https://zenodo.org/records/10925359} and the development server is hosted by Github at \url{https://www.github.com/Stew-McD/T-reX}. A user guide with comprehensive documentation are presented at \url{https://T-reX.readthedocs.io}.

@@

\section*{CRediT authorship contribution statement} % VISUAL VERSION

>>>>

%% 
%% Copyright 2019-2024 Elsevier Ltd
%% 
%% This file is part of the 'CAS Bundle'.
%% --------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.3c of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.3c or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%%
%
% T-reX manuscript using the elsevier template for JRC
%
% Repository for the T-reX manuscript at https://github.com/Stew-McD/T-reX_Publication
% Repository for the T-reX python program at https://github.com/Stew-McD/T-reX
%
% S.C.McDowall - github/Stew-McD
%
%%%%

%% 
%% The list of all files belonging to the 'CAS Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for cas-dc documentclass for 
%% double column output.

%\documentclass[a4paper,fleqn]{cas-dc}

@@

\documentclass[a4paper,fleqn,longmktitle]{cas-dc}

\usepackage[numbers]{natbib}
%\usepackage[authoryear]{natbib}
%\usepackage[authoryear,longnamesfirst]{natbib}

@@

@@

% Abstract\\
\begin{abstract}

The quintessential principle of the circular economy is to minimise material consumption and waste generation. Thus, identifying and quantifying waste and material flows is critical. Life Cycle Assessment (LCA) is a powerful quantitative method for this, especially given its capacity to expose the details of an activity's entire life cycle, often guiding the implementation of circular principles to where they could be most effective.

Pursuant to these goals the authors have developed 'T-reX', an analytical tool for LCA, written in Python. T-reX extends on the existing computational ecosystem centered on Brightway and Premise and can facilitate the quantification of user-defined supply chain demands for activities in current and prospective scenarios. T-reX streamlines database manipulation for LCA practitioners, integrating methods to aggregate and analyse demand inventories that could expose hotspots of hidden risk.

With a simple case study using Lithium-ion batteries, we demonstrate some of T-reX's potential by detailling and exploring aspects of their waste and material footprints. Since such footprints are often linked to negative externalities, T-reX could support sustainable decision-making and contribute to the development of the circular economy by promoting the understanding of our material consumption and waste generation.

@@

% Keywords
\begin{keywords} 
circular economy \sep waste \sep material \sep life cycle assessment \sep critical raw material \sep supply chain
\end{keywords}

% Graphical abstract
\begin{graphicalabstract}
  \includegraphics{figs/T-reX_graphical-abstract.png}
\end{graphicalabstract}

\clearpage

%% GRAPHICAL ABSTRACT FULL SIZE
\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{figs/T-reX_graphical-abstract.png}
	\caption{The graphical abstract of the T-reX program developed for waste and material footprinting in Life Cycle Assessment}
	\label{fig:graphical_abstract}
\end{figure}

\clearpage

% Switch to landscape for the included PDF
\newgeometry{landscape}
\includepdf[pages=1, width=\textheight]{figs/T-reX_graphical-abstract.pdf}
\restoregeometry % Restore the original geometry

@@

@@

The development of a circular economy has become a central focus in the pursuit of sustainability objectives toward curtailment of our environmental footprint within planetary boundaries~\citep{eu2019greendeal, eu2020circ,nl2023ceplan,nl2016ceplan,pardo2018ce,ellenmacarthur2015ce}. Fundamental to this development is a decrease in primary material consumption and a reduction of life cycle waste through the implementation of `re-X' strategies (e.g., refuse, rethink, design for---and implementation of---repair, remanufacturing and recycling)~\citep{reike2018rex, eu2022ecodesign, eu2022repair, eu2015reman}. In addition to circular economy goals, contemporary geopolitical tensions in an ever more globalised world have highlighted the vulnerability of many advanced economies to intentional supply disruptions---wrought as an act of competition or even outright hostility~\citep{jrc2023supplychain,hartley2024cepolitics,berry2023crm}.

@@

Life Cycle Assessment (LCA) is a useful method for the holistic estimation of the environmental impacts of products and processes. LCA can comprehensively evaluate impacts across the entire life cycle---from `cradle to grave'--- often identifying critical hotspots and guiding prioritisation of action~\citep{guinee2011lca}. The standard approach is to apply Life Cycle Impact Assessment (LCIA) methods (such as ReCiPe~\citep{huijbregts2016recipe} and CML~\citep{guinee2002cml}), which convert the inventory data into a set of impact scores based on the sum of the elementary flows (those between the technosphere and biosphere). These scores can then be aggregated into one metric for each impact category that can be compared across products and processes. Extending on standard LCA is \textit{ex-ante} LCA, which employs future scenarios to construct `prospective background databases' to predict the impacts of supply chains that have not yet been (and may never be) realised~\citep{cucurachi2018exante,blanco2020exante}.

@@

Following the logic of the standard LCA models \cite{ecoinvent2016version3, guinee2004economicallocation}, the identification of material exchanges with T-reX differs from that used to identify waste exchanges. That is, the search queries are based on the names of the relevant `market activities' for the material of interest rather than keywords in the exchange's name. A useful feature of T-reX is that, in cases where there are several markets for one material or material grouping, the program can aggregate these flows. For example, exchanges with markets for the rare-earth-elements (REEs) `market for cerium', `market for dysprosium', `market for erbium', etc.\ can be aggregated into a single indicator category for REEs. Similarly, the total demand for all critical raw materials (CRMs) can be calculated in the same manner.

@@

\subsubsection{LCA calculations}
For each combination of activity, method, and database, a total summation was calculated along with details of the top contributing processes. Additionally, for the T-reX methods, a contribution analysis was performed with the \texttt{bwa.compare\_activities\_by\_grouped\_leaves} function from the \texttt{brightway2\_analyzer} package~\citep{mutel2016brightway2analyzer}, a component of the \texttt{brightway} ecosystem. This function performs one subjective method of graph traversal on the impact matrix of the LCA object to a specified cutoff. The groups of the resulting 'leaves' are collected by their Cooperative Patent Classification (CPC) codes [giving a representation of industry sector involvement]. T-reX provides, essentially, a tabulation of internal exchanges and an insight into the products and sectors in the supply chain of the activity that carry the most substantial responsibility for the final waste generation or material demand footprints.

@@

% Back matter
\section*{Data availability}
All data used is publicly available online under the noted sources. T-reX is fully open source and dedicated to the public domain under the CC-1.0 licence. It is installable via the Python Package Index (PyPI) under the name `T\_reX\_LCA' \url{https://pypi.org/project/T_reX_LCA}.
The full source code for T-reX is archived by Zenodo at \url{https://zenodo.org/records/10925359} and the development server is hosted by Github at \url{https://www.github.com/Stew-McD/T-reX}. A user guide with comprehensive documentation are presented at \url{https://T-reX.readthedocs.io}.

@@

\section*{CRediT authorship contribution visualisation} % IMPROVED VISUAL VERSION
